---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hydrodownloadR

<!-- badges: start -->
[![R-CMD-check](https://github.com/bafg-bund/hydrodownloadR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/bafg-bund/hydrodownloadR/actions/workflows/R-CMD-check.yaml)
[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](LICENSE)
[![lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

**Acknowledgements.** The repository structure is inspired by Ryan Riggs’
**RivRetrieve**. Thanks to **Frederik Kratzert** (co-author of the referenced
publication) for helpful exchanges and for maintaining and porting the
RivRetrieve concept to Python in **RivRetrieve-Python**.

**hydrodownloadR** provides a unified, extensible interface for discovering
hydrological stations and downloading daily time series (e.g., water discharge,
water level, water temperature, and several other water quality parameters) from
national/regional public APIs. The package uses a provider registry with
identifiers (`provider_id`, `provider_name`) and S3 generics `stations()` /
`timeseries()`. It supports complete histories (1900-01-01 until today),
per-station selection, rate limiting and retries, optional authentication via
environment variables, UTF-8 to ASCII normalization, and coordinate
transformation to WGS84.

> Built for reproducible workflows and straightforward addition of new providers.

------------------------------------------------------------------------

## Why this package exists

This package is developed and used at the Global Runoff Data Centre (GRDC, BfG)
as part of reproducible workflows to discover stations and retrieve update time
series from public APIs. It is also suitable for other global data centres
hosted by BfG (Federal Institute of Hydrology) that rely on consistent, auditable
data access and update pipelines.

------------------------------------------------------------------------

## Installation

### Development version (GitHub)

```{r, eval = FALSE}
# Option A: remotes
install.packages("remotes")
remotes::install_github("bafg-bund/hydrodownloadR")

# Option B: pak (fast)
install.packages("pak")
pak::pak("bafg-bund/hydrodownloadR")
```

### CRAN version (planned)

```{r, eval = FALSE}
install.packages("hydrodownloadR")
```

------------------------------------------------------------------------

## Quick start

List available providers:

```{r, eval = FALSE}
library(hydrodownloadR)

hs <- hydro_services()
hs
```

`hs` includes an overview of providers and (where available) licensing / terms
information to help users understand access and reuse conditions.

Select one provider and list stations:

```{r, eval = FALSE}
x <- hydro_service(hs$provider_id[1])
stn <- stations(x)
stn
```

Check which time series parameters are available for this provider:

```{r, eval = FALSE}
timeseries_parameters(x)
```

Download a daily time series for one station:

```{r, eval = FALSE}
ts <- timeseries(
  x,
  stations  = stn$station_id[1],
  parameter = "water_discharge",
  mode      = "complete"
)
ts
```

------------------------------------------------------------------------

## Provider-specific options and "complete history"

Upstream APIs differ. Some providers expose additional options such as
authentication, quality flags, paging strategies, or "complete history" modes.
These options may evolve as providers change their APIs. For best results:

- Start with defaults.
- Check the help pages: `?hydro_service`, `?stations`, `?timeseries`,
  `?timeseries_parameters`.
- Review provider-specific notes (when linked from the provider registry).

Tip: if an adapter exposes a flag that must be explicitly enabled (for example,
a complete-history mode), set it explicitly in your scripts so results remain
reproducible and robust to upstream changes.

------------------------------------------------------------------------

## Licensing and terms of use

This package provides technical access to data services. Data access and reuse
are governed by each provider's terms and licensing conditions. Please review
and comply with those terms and licensing conditions when using downloaded data.

------------------------------------------------------------------------

## Contributing and support

Contributions are welcome:

- Use GitHub Issues for bug reports, feature requests, and new provider requests.
- If you report a bug, please include provider_id, station_id (if relevant),
  parameter, time range, a minimal reproducible snippet, and `sessionInfo()`.

Pull requests are welcome. Please ensure CI is green (R CMD check) and keep new
code ASCII-friendly where possible.

------------------------------------------------------------------------

## Development note (AI assistance)

Portions of this project were drafted with AI assistance (ChatGPT) and then
reviewed, edited, and tested by the maintainer. All changes are maintained via
version control and validated with automated checks.
